Viscous Erosion in Porous Media

### 2018-03-27 docs/methods/erosion_methods.tex
NM

The corresponding stream function is biharmonic, $\Delta^2 \psi = 0$, and the ansatz $\psi = r^{\lambda}f(\theta)$ gives the 4th-order ODE~\cite{poz1997}
\begin{align}
\label{fODE}
f'''' + 2(\lambda^2 - 2 \lambda + 2)f'' + \lambda^2(\lambda-2)^2 f = 0.
\end{align}
This ODE is subject to no-slip boundary conditions along the wedge $f(\thb) = f'(\thb) = 0$ and odd symmetry about the horizontal axis, $f(0) = f''(0) = 0$. For arbitrary opening angle, ODE \eqref{fODE} and the boundary conditions comprise an eigenvalue problem for $\lambda$, whose solution furnishes the stream function and the corresponding shear stress on the wedge surface $\tau = \mu r^{\lambda-2} f''(\thb)$. We, however, seek to determine the particular opening angle that produces {\em uniform} shear stress and therefore set $\lambda = 2$.
The odd-symmetric solution to \eqref{fODE} that corresponds to $\lambda=2$ is given by
\begin{align}
  f(\theta) = B \sin (2 \theta) + C \theta \, .
\end{align}
Imposing no slip along the wedge surface gives the condition
\begin{align}
  \sin(2 \thb) - 2 \thb \cos(2 \thb) = 0 \, .
\end{align}
This nonlinear equation has solution $\thb \approx 129^\circ$, which implies that the the uniform-stress wedge has an opening angle of $\oangle \approx 102^\circ$.
% Pozrikidis page 239: symmetric flow.

@@

We now aim to measure the opening angles formed in our numerical simulations to compare against this prediction. However, Section \ref{sec:scaling} shows that the magnitude of the shear stress diverges as the body vanishes, which causes our simulations to lose accuracy in the final, critical moments when the body is expected to approach its limiting shape. To circumvent this problem, we remove the effect of shrinking by forcing $\Vn$ to have zero mean,
\begin{align}
\Vn \leftarrow \Vn - \mean{\Vn} \, .
\end{align}
This modification allows erosion to sculpt the shape as it would normally, but with the area artificially held fixed.

@@

	We show in Fig.~\ref{fixed_intface}(a) several interfaces of this modified erosion process, with the same spatial resolution, $N_\iin = 1024$, as before. Notice the body obtains a similar profile as observed previously, but with fixed area. Importantly, more numerical accuracy is now available to quantify the terminal state. The inset in Fig.~\ref{fixed_intface}(a) is a zoom showing that, due to our numerical regularization, the near corner at the body's rear is actually smooth on a fine scale. We show in Fig.~\ref{fixed_intface}(b) the stress distributions on the same interfaces. As observed before, the stress becomes more evenly distributed as erosion reshapes the body. The slight rise in shear near the front/rear corners that was observed in Fig.~\ref{shrink_intface}(b) is no longer present, suggesting that this effect was due to loss of accuracy. Rather, the late-time stress distributions in Fig.~\ref{fixed_intface}(b) decrease monotonically to zero near the stagnation points. In this figure, time is normalized by the same value $t_f = 1.79 \times 10^{-2}$ as in Fig.~\ref{shrink_intface}.

>>>>

The corresponding stream function is biharmonic, $\Delta^2 \psi = 0$, and the ansatz $\psi = r^{\lambda}f(\theta)$ gives the 4th-order ODE~\cite{poz1997}
\begin{align}
\label{fODE}
f'''' + 2(\lambda^2 - 2 \lambda + 2)f'' + \lambda^2(\lambda-2)^2 f = 0.
\end{align}
This ODE is subject to no-slip boundary conditions along the wedge $f(\thb) = f'(\thb) = 0$ and odd symmetry about the horizontal axis, $f(0) = f''(0) = 0$. For arbitrary opening angle, ODE \eqref{fODE} and the boundary conditions comprise an eigenvalue problem for $\lambda$, whose solution furnishes the stream function and the corresponding shear stress on the wedge surface $\tau = \mu r^{\lambda-2} f''(\thb)$. We, however, seek to determine the particular opening angle that produces {\em uniform} shear stress and therefore set $\lambda = 2$.
The odd-symmetric solution to \eqref{fODE} corresponding to $\lambda=2$ is
\begin{align}
  f(\theta) = B \sin (2 \theta) + C \theta \, .
\end{align}
Imposing no slip along the wedge surface gives the condition
\begin{align}
  \sin(2 \thb) - 2 \thb \cos(2 \thb) = 0 \, .
\end{align}
This nonlinear equation has solution $\thb \approx 129^\circ$, which implies that the the uniform-stress wedge has an opening angle of $\oangle \approx 102^\circ$.
% Pozrikidis page 239: symmetric flow.

@@

We now aim to measure the opening angles formed in the numerical simulations to compare against the above prediction. However, Section \ref{sec:scaling} shows that the magnitude of the shear stress diverges as the body vanishes, which causes our simulations to lose accuracy in the final, critical moments when the body is expected to approach its limiting shape. To circumvent this problem, we remove the effect of shrinking by forcing $\Vn$ to have zero mean,
\begin{align}
\Vn \leftarrow \Vn - \mean{\Vn} \, .
\end{align}
This modification allows erosion to sculpt the shape as it would normally, but with the area artificially held fixed.

@@

	We show in Fig.~\ref{fixed_intface}(a) several interfaces of this modified erosion process, with the same spatial resolution, $N_\iin = 1024$, as before. Notice the body obtains a similar profile as observed previously, but with fixed area. Importantly, more numerical accuracy is now available to quantify the terminal state. The inset in Fig.~\ref{fixed_intface}(a) zooms on the rear of the body to demonstrate that, due to our numerical regularization, what appears to be a corner is actually smooth on a fine scale. We show in Fig.~\ref{fixed_intface}(b) the stress distributions on the same interfaces. As observed before, the stress becomes more evenly distributed as erosion reshapes the body. The slight rise in shear near the front/rear corners that was observed in Fig.~\ref{shrink_intface}(b) is no longer present, suggesting that this effect was due to loss of accuracy. Rather, the late-time stress distributions in Fig.~\ref{fixed_intface}(b) decrease monotonically to zero near the stagnation points. In this figure, time is normalized by the same value $t_f = 1.79 \times 10^{-2}$ as in Fig.~\ref{shrink_intface}.
